---
title: "Simulation model"
author: "Mikaela Provost"
date: "Sept 19, 2017"
output: html_document
---

Get my packages:
```{r}
require(ggplot2)
require(tidyr)
#install.packages("Reunran","foreach")
#library(Runuran)
require(foreach)
```

```{r}
# load parameters & functions
#source("C:/Users/provo/Documents/GitHub/popdy/popdy_code/parameters_pacifichake.r")
source("C:/Users/provo/Documents/GitHub/popdy/popdy_code/parameters_pacificmackerel.r")
source("C:/Users/provo/Documents/GitHub/popdy/popdy_code/functions.r")
# note: in 'functions' this includes - 
# 1) assemble_leslie
# 2) is_mature
```

```{r}
# length of simulation in years
time = 5100

# if want recruitment variability, define eta as a normal random variate 
# for larval mortality, # producing lognormal survival deviations
eta = rnorm(n = time, mean = 0, sd = eta_sd) - ((eta_sd ^ 2) / 2)

# if you want M deterministic
M = rep(natural_mort, time)

# WRITE FUNCTION TO SET UP STRUCTURES FOR STORING DATA IN THE CORRECT 
# FORMAT, GIVEN SCENARIO ARGUMENTS

# set up empty matrix for size at age at time
W = matrix(0, a_max, time)
# If deterministic growth, define W as same accross all times 
# (FOR PERFORMANCE, IN FUTURE DON'T DEFINE FOR EACH TIME)
W[,] = W_inf * (1 - exp(-k * ((1:a_max) - t_0))) ^ b

# set up empty matrix for maturity at age, given size, at time
mat = matrix(0, a_max, time)
```

If deterministic growth, fill the mat matrix with constant values for maturity at age, based on deterministic growth.

plogis() gives the distribution function for the logistic distribution with parameters 'location' and 'scale'. The output from plogis() gives the prediction at age and time for maturity. 

Right now it's constant across all ages, this seems wrong (?)

location parameter = mean, W_mat=size at 50% mat
scale parameter = variance, standard deviation * (pi^2/3)
see R help pg: https://stat.ethz.ch/R-manual/R-devel/library/stats/html/Logistic.html

```{r}
mat[,] <- plogis(W[,1], location = W_mat, scale =  W_mat_sd * (sqrt(3)/pi))

# set up empty matrix for abundance at age at time
n = matrix(NA, a_max, time)

# set up empty vector to store total egg production each year
eggs = rep(NA, time)
# set up empty vector to store catch each year
catch = rep(NA, time)

# keep seed constant for repeatability initially
seed <- 42
set.seed(seed)

# package parms with initial values for some variables, to pass to top-level function
parms_plus = c(eta = eta, M = M, time = time, W = W, mat = mat, n = n, eggs = eggs,
               a_max = a_max, k = k, W_inf = W_inf, b = b, L_W_slope = L_W_slope, catch = catch,
               W_inf = W_inf, t_0 = t_0, W_mat = W_mat, W_mat_sd = W_mat_sd, W_rec = W_rec,
               W_rec_sd = W_rec_sd, W_esc = W_esc, W_esc_sd = W_esc_sd, f_W_slope = f_W_slope,
               f_W_int = f_W_int, alpha = alpha, beta = beta, #W_inf_sd = W_inf_sd, 
               #W_inf_min = W_inf_min, W_inf_max = W_inf_max, 
               R_0 = R_0, natural_mort = natural_mort)

```

Simulation function
Write function to simulate, given parms and value for harvest rate (H), noting that for Pacific Hake the historical range is 0.05 to 0.4.
```{r}
simulate <- function(parms_plus = parms_plus, H = 0) {
  
  # fill in abundance at age for time = 1 given that pop is 
  # fished at fishing mortality H at all ages and at the SAD
  n[,1] = R_0 * exp(- (natural_mort + H) * (0:(a_max-1)))
  
  for(j in 2:time){ # step through each time step, 
                    # now that we have abundance at age in t=1
    
    # at this time step, calculate the Leslie matrix, total eggs, total cal
    leslie_eggs_catch <- assemble_leslie(SR = 'Beverton_Holt', mat_a_array = mat, 
                                         W_a_array = W, year = j, n = n, H = H, 
                                         W_rec = W_rec, W_rec_sd = W_rec_sd,                                                                            W_esc = W_esc, W_esc_sd = W_esc_sd, M = M,
                                         f_W_slope = f_W_slope, alpha = alpha, 
                                         beta = beta, eta = eta)
    # extract Leslie matrix for time j
    A <- leslie_eggs_catch$A
    # store egg production for time j
    eggs[j] <- leslie_eggs_catch$E
    # store catch for time j
    catch[j] <- leslie_eggs_catch$C
  
    # perform population projection of one time step,
    # gets abundance at age in t+1
    n[,j] <- A %*% n[,j-1]
  
    # make sure abundances at age that are less than 1 become zero
    for(i in 1:a_max) { if(n[i,j] < 1) n[i,j] = 0 }
    # if whole population goes below 1, stop the simulation
    if(sum(n[,j]) < 1) print("Population crashed!") 
  }

  # Return list including numbers at age, egg production, 
  # and catch for the whole timeseries, with first 100 time 
  # steps trimmed off (to make sure we are around the equilibrium)

  # n = a matrix with a_max rows, and time-100 columns
  # eggs = a vector, egg production each yr
  # catch = a vector, catch each yr
return(list(n=n, eggs=eggs, catch=catch))
}
```

For a range of harvest levels, generate output (n, eggs, catch). This ouptut is used to generate timeseries plots of eggs, abundance, catch over time. And these are used to calculate specturm analysis. 
```{r}
#output = foreach(H. = seq(0, 2.4, 0.4)) %do% simulate(H = H.)
output = simulate(H=0)
plot(output$eggs)
plot(output$n)


```


Ignore this section (stochastic growth)
```{r}

# Else stochastic growth
# if updating W_inf within function, just insert deterministic size at age into first column of empty size at age over time matrix
#W[,1] = W_inf * (1 - exp(-k * ((1:a_max) - t_0))) ^ b
# else calculate random series of W_inf and use to preallocate series of size at age over time
#W_inf = urlnorm(time, meanlog = log(W_inf), sdlog = log(W_inf_sd), lb = 0, ub = W_inf_max - W_inf_min) + W_inf_min
# NEW GROWTH FORM NOT WORKING, USE LENGTH-BASED VERSION BELOW
#for(j in 2:time) {
#  W[1,j] =  W_inf[j-1] * (1 - exp(-k * (1 - t_0))) #^ b #W_0 + (W_inf[j-1] * exp(-b * k * (1-t_0)) * (1 - exp(-b * k)))
#  W[2:a_max,j] = W[1:(a_max-1),j-1] + W_inf[j-1] * exp(-k * ((1:(a_max-1)) - t_0)) * (1 - exp(-k)) 
#}
```
