---
title: ""
author: "Mikaela Provost"
date: "9/26/2018"
output:
  html_document: default
  pdf_document: default
---

# This code creates grid plots: spectral analysis at different k values for each cod pop

### Plan
1. read in LSBlist = matricies of LSB at age for diff k, for each pop
2. use LSBlist to generate Jacobian matricies for simulations
3. get output from simulations ready for plotting
4. plot 1 - 4 panels, all pops together
5. plot 2 - 4 rows (k vals), columns are different peak spawning ages

```{r, include=FALSE}
# --
# load functions, libraries
library(ggplot2)
library(gridExtra)
library(dplyr)
library(plyr)
library(ggrepel)
library(devtools)
library(broom)
source("C:/Users/provo/Documents/GitHub/popdy/cod_code/2_cod_functions.r") # load functions
source("C:/Users/provo/Documents/GitHub/popdy/cod_code/0_load_cod_data.r") # load cod data
source("C:/Users/provo/Documents/GitHub/popdy/cod_code/simulation_model_cod_v3.r") # load sim model, set up to use BH but can feed in Jacobian or Leslie

# read in eigentable - I'm using some information from the table
eigentable <- read.csv(file="C:/Users/provo/Documents/GitHub/popdy/cod_code/mikaelaLSB/eigentable.csv",header=T)
eigentable = as.data.frame(eigentable)
```

```{r, warning=FALSE}
# ---
F.halfmax = 0 #for now, F is 0
tknot =0
ks = c(0.5,1) #set k values

# ---
# Create Jacobian for each pop
# first, calculate LSB at age. Values of k slope are NOT incorporated here
LSBlist <- as.list(rep(NA,length(codNames))) #LSB at age for pops stored here
names(LSBlist) <- codNames

for (i in 1:length(LSBlist)) { # for each population
  # load parms for cod pop i
  source(file = paste('C:/Users/provo/Documents/GitHub/popdy/cod_pops/',names(LSBlist)[i], '.r', sep=''))
  # this should load parms: L_inf, K, TEMP, maxage, B0, B1, tknot=0
  
  # for each pop calc LSB at age (multiply by k in next loop, generate matrix)
  #lsb.at.k = matrix(0,nrow=maxage,ncol=1) #create empty matrix to store LSB at age
  lsb = calculate_LSB_at_age_by_F(data=datalist[[i]], maxage=maxage,
                                             L_inf=L_inf, K=K, TEMP=TEMP, F.halfmax=F.halfmax,
                                             B0=B0,B1=B1)
  LSBlist[[i]] = lsb #LSBlist should have length of codNames
  # LSB = wt at age * prop mature at age * survival at age
}
rm(i,lsb)
```

## Two options: use (1) rescaled Leslie or (2) Leslie matrix in simulations. The first chunk of code is for using the rescaled Leslie, or run the section block of code to use the Leslie. Need to rename one of them as A3dlist to move on.

### Goal: create list of arrays with re-scaled Leslie matricies for each k-population
```{r,warning=FALSE}
Ls3dlist = as.list(rep(NA,length(LSBlist))) #store rescaled Leslie arrays for each pop here
names(Ls3dlist) = names(LSBlist)

for (i in 1:length(LSBlist)) { #for each pop i
  lsb = LSBlist[[i]] #extract LSB vectors
  Lslist = as.list(rep(NA,length(ks))) #create a list for diff k values
  
  for (j in 1:length(ks)) { #this assumes I have LSB at the desired k values
    # start with empty matrix
    A = matrix(0,nrow=length(lsb),ncol=length(lsb)) #create an empty matrix
    
    # create new rescaled Leslie matricies at each k
    #A[1,]=lsb/sum(lsb) #fill in relative LSB at age across top row
    A[1,]=lsb
    A[1,]=A[1,]*ks[j] #multiply relative fecundities by k
    
    for(a in 2:length(A[,1])-1){ #filling in survivals
      A[a+1,a]=1} #fill in survival of 1 on subdiagonal
    
    Lslist[[j]] = A
    
  }
  Ls3d = array(unlist(Lslist),dim=c(length(A[,1]),length(A[,1]),length(ks)))
  Ls3dlist[[i]] = Ls3d #store array 
}
rm(i,j,a,Lslist,Ls3d)

# reassign Ls3dlist to A3dlist
A3dlist = Ls3dlist
```

### Run this section if you want to use *Leslie* matricies and not Jacobian matricies
```{r}
A3dlist = as.list(rep(NA,length(LSBlist))) #store Leslie arrays for each pop here
names(A3dlist) = codNames

for (i in 1:length(A3dlist)) { # for each population
  # load each Leslie matrix
  A <- read.table(file = paste('C:/Users/provo/Documents/GitHub/popdy/cod_code/mikaelaLeslie/matrix_maxages/',
                                                               names(A3dlist)[i], '.txt', sep=''))
  Alist = as.list(rep(NA,length(ks))) #create a list for diff k values
  
  for (j in 1:length(ks)) { #this assumes I have LSB at the desired k values
    # start with empty matrix
    A[1,] = A[1,] * ks[j]
    Alist[[j]] = A
  }
  A3d = array(unlist(Alist),dim=c(length(A[,1]),length(A[,1]),length(ks)))
  A3dlist[[i]] = A3d #store array 
}
rm(i,j,A,Alist,A3d)

```

```{r}

timesteps = 1000
rm_first_timesteps = 100
alpha = 100
beta = 1000
initial_eggs = 1000
sig_r = 0.1
span.multiplier = 1 # what is this again?
```

### The code below uses Jacobian or Leslie matricies associated with each population and k slope to simulate timeseries of egg, recruits, and spawning adults. The matricies are stored in a list of 3D arrays.
```{r}
freq = seq(from=0.001111111, to=0.5, by=0.001111111) #all frequencies, for plotting later
output.norm.3dList = as.list(rep(NA,length(A3dlist)))

raw.3dList = as.list(rep(NA,length(A3dlist)))
names(raw.3dList) = names(A3dlist)

# setting 'span' - a vector of odd integers to specify the smoothers
tmp <- ceiling(sqrt(length(1:(timesteps-rm_first_timesteps-1)))) #square root of timeseries length, rounded
if (tmp %% 2 == 0) {m <- tmp+1} else {m <- tmp} #make it odd, if the square root is even
m = m * span.multiplier
    
for (y in 1:length(A3dlist)) { #for pop y
  A3d = A3dlist[[y]] #choose the 3d array of matricies, use these to simulate below
  
  outputL.raw = as.list(rep(NA,length(ks))) #create temp empty list for storing output at each k 
  outputL.spec = as.list(rep(NA,length(ks)))#create temp empty list for storing output at each k
                         
  for (k in 1:length(ks)) { #for each k value
    
    A = A3d[,,k] #choose the Leslie matrix for this k value
   
    output = sim_model(A=A, timesteps=timesteps, #simulate
                       alpha=alpha, beta=beta, 
                       sig_r=sig_r, initial_eggs=initial_eggs)  
    outputL.raw[[k]] = do.call(cbind,output[c(2,3)]) #save sim output for each pop 
    
    # --- do the spectral analysis
    spec_all = matrix(NA,nrow=length(freq),ncol=2) # matrix to store spec of egg, recruits
    
    egg.sim = output[[2]][rm_first_timesteps:(timesteps-2)] #extract simulation of egg timeseries from model output
    spec_egg = spec.pgram(x=egg.sim, 
                          spans=c(m,m),plot = FALSE)
    spec_all[,1] <- spec_egg$spec #save spec output for plotting pops together, Helen Wearing says to multiply by 2
    
    rec.sim = output[[3]][rm_first_timesteps:(timesteps-2)] #extract sim of recruits
    spec_rec = spec.pgram(x=rec.sim, 
                          spans=c(m,m),plot = FALSE)
    spec_all[,2] = spec_rec$spec #save spec output for plotting pops together, Helen Wearing says to multiply by 2
    
    #adu.sim = output[[4]][rm_first_timesteps:(timesteps-2)] #extract sim of spawning adults
    #spec_adu = spec.pgram(x=adu.sim, 
    #                      spans=c(m,m),plot = FALSE)
    #spec_all[3,] = spec_adu$spec #save spec output for plotting pops together, Helen Wearing says to multiply by 2
    
    outputL.spec[[k]] = spec_all
  }
  output.norm.3d = array(unlist(outputL.spec),dim=c(length(freq),2,length(ks)))
  output.norm.3dList[[y]] = output.norm.3d #store array 
  output.raw.3d = array(unlist(outputL.raw), dim=c(length(outputL.raw[[1]][,1]),2,length(ks)))
  raw.3dList[[y]] = output.raw.3d
}

# convert this into a dataframe
names(output.norm.3dList) = names(A3dlist)
# for each pop, 3d object: 3 rows=eggs/rec/adu, col=spec output, stack=diff k 
```

### Try plotting the timeseries at different k values for each population. 
```{r}
rawpoplist = as.list(rep(NA,length(raw.3dList)))
names(rawpoplist) <- names(raw.3dList)

for (e in 1:length(raw.3dList)) { #for each pop
  a = raw.3dList[[e]] #convert one pop array to df at a time
  
  rawlist = as.list(rep(NA,length(ks))) #placeholder for storing spec output at k
  
  for (k in 1:length(ks)) { #for each k, convert sim output into df
  
    rawout = as.data.frame(a[,,k]) #choose simout for each k
    names(rawout) = c("eggs","recruits")
    rawout$timesteps = seq(from=1,to=length(rawout$eggs),by=1)
    rawlong = melt(rawout,id.vars="timesteps")
    names(rawlong) <- c("timesteps","out.type","value")
    rawlong$kval <- rep(ks[k],length=length(rawlong$timesteps))
    rawlist[[k]] = rawlong #store sim output df for each k
  } 
  df <- ldply(rawlist, data.frame)
  df$codNames <- rep(codNames[e],length=length(df$timesteps))
  rawpoplist[[e]] = df
}

raw.out.df <- ldply(rawpoplist, data.frame)
raw.out.df$kval <- factor(raw.out.df$kval)
raw.out.df$codNames <- factor(raw.out.df$codNames)
p <- ggplot(raw.out.df[raw.out.df$kval %in% c(0.2,0.5,0.9),],aes(x=timesteps,y=value,group=kval)) +
  geom_point(aes(color=kval)) +
  facet_grid(.~codNames)
pdf(file='C:/Users/provo/Documents/GitHub/popdy/cod_figures/timeseries_plots_for_kval_JacBH.pdf', width=30, height=6)
p
dev.off()
```

### The spec output from above is stored in a list of 3D arrays. The code below converts this list into a df.
```{r}
simpoplist = as.list(rep(NA,length(output.norm.3dList)))
names(simpoplist) <- names(output.norm.3dList)

for (e in 1:length(output.norm.3dList)) { #for each pop
  a = output.norm.3dList[[e]] #convert one pop array to df at a time
  
  simlist = as.list(rep(NA,length(ks))) #placeholder for storing spec output at k
  
  for (k in 1:length(ks)) { #for each k, convert sim output into df
  
    simout = as.data.frame(a[,,k]) #choose simout for each k
    names(simout) = c("eggs","recruits")
    simout$freq = freq
    simlong = melt(simout,id.vars="freq")
    names(simlong) <- c("freq","out.type","value")
    simlong$kval <- rep(ks[k],length=length(freq))
    simlist[[k]] = simlong #store sim output df for each k
  } 
  df <- ldply(simlist, data.frame)
  df$codNames <- rep(codNames[e],length=length(df$freq))
  simpoplist[[e]] = df
}

spec.out.df <- ldply(simpoplist, data.frame)
table(spec.out.df$codNames)
```


### Calculating the AUC.
```{r}
head(spec.out.df)
test <- spec.out.df[spec.out.df$out.type == "eggs" &
            spec.out.df$kval == "0.2" &
            spec.out.df$codNames == "Northsea",]
ggplot(test, aes(x=freq,y=value)) + geom_point()
sum(test$value * min(test$freq))

# Eggs
aucList <- as.list(NA,length=length(codNames))
for (i in 1:length(codNames)) {
  d <- spec.out.df[spec.out.df$codNames == codNames[i] & spec.out.df$out.type == "eggs",]
  auc <- rep(NA,length=length(ks))
  
  for (k in 1:length(ks)) {
    auc[k] <- sum( d[d$kval == ks[k],]$value * min(d$freq) )}
  
  aucdf <- data.frame(cbind(auc,ks,rep(codNames[i],length=length(ks))))
  names(aucdf) <- c("auc","kval","codNames")
  aucList[[i]] <- aucdf
}
auc.df.eggs <- ldply(aucList, data.frame)
rm(i,k,d,auc,aucdf,aucList)

# Recruits
aucList <- as.list(NA,length=length(codNames))
for (i in 1:length(codNames)) {
  d <- spec.out.df[spec.out.df$codNames == codNames[i] & spec.out.df$out.type == "recruits",]
  auc <- rep(NA,length=length(ks))
  
  for (k in 1:length(ks)) {
    auc[k] <- sum( d[d$kval == ks[k],]$value * min(d$freq) )}
  
  aucdf <- data.frame(cbind(auc,ks,rep(codNames[i],length=length(ks))))
  names(aucdf) <- c("auc","kval","codNames")
  aucList[[i]] <- aucdf
}
auc.df.recruits <- ldply(aucList, data.frame)
rm(i,k,d,auc,aucdf,aucList)

# Spawning adults
aucList <- as.list(NA,length=length(codNames))
for (i in 1:length(codNames)) {
  d <- spec.out.df[spec.out.df$codNames == codNames[i] & spec.out.df$out.type == "Nsize",]
  auc <- rep(NA,length=length(ks))
  
  for (k in 1:length(ks)) {
    auc[k] <- sum( d[d$kval == ks[k],]$value * min(d$freq) )}
  
  aucdf <- data.frame(cbind(auc,ks,rep(codNames[i],length=length(ks))))
  names(aucdf) <- c("auc","kval","codNames")
  aucList[[i]] <- aucdf
}
auc.df.adu <- ldply(aucList, data.frame)
rm(i,k,d,auc,aucdf,aucList)

```
### Ploting
```{r}

spec.out.df <- merge(spec.out.df,eigentable[c("codNames","temp")],all.x=TRUE)
spec.out.df <- merge(spec.out.df,eigentable[c("codNames","mode_age")],all.x=TRUE)
spec.out.df <- merge(spec.out.df,eigentable[c("codNames","sd_mode")],all.x=TRUE)
spec.out.df <- merge(spec.out.df,eigentable[c("codNames","max_ages")],all.x=TRUE)
spec.out.df <- merge(spec.out.df,eigentable[c("codNames","cvs_mode")],all.x=TRUE)


# Plot spectra
recruit.df <- spec.out.df[spec.out.df$out.type=="recruits" &
                            spec.out.df$kval %in% c("0.2","0.5","0.8"),]
recruit.df <- mutate(recruit.df, codNames = reorder(codNames, mode_age))
rp<- ggplot(recruit.df,
       aes(x=freq,y=value)) +
  geom_line() +
  facet_grid(kval~codNames) + ggtitle("recruits: increasing in peak spawning age")

# ---
egg.df <- spec.out.df[spec.out.df$out.type=="eggs" &
                            spec.out.df$kval %in% c("0.2","0.5","0.8"),]
egg.df <- mutate(egg.df, codNames = reorder(codNames, mode_age))
ep <- ggplot(egg.df,
       aes(x=freq,y=value)) +
  geom_line() +
  facet_grid(kval~codNames) + ggtitle("eggs: increasing in peak spawning age")
# ---
adu.df <- spec.out.df[spec.out.df$out.type=="adults" &
                            spec.out.df$kval %in% c("0.2","0.5","0.8"),]
ap <- ggplot(adu.df,
       aes(x=freq,y=value)) +
  geom_line() +
  facet_grid(kval~codNames) + ggtitle("spawning biomass")

#pdf(file='C:/Users/provo/Documents/GitHub/popdy/cod_figures/spectrum_plots_v1.pdf', width=20, height=7)
#rp
#ep
#dev.off()





```