# This code creates grid plots: spectral analysis at different k values
# for each cod pop

# Plan
# 1. read in LSBlist = matricies of LSB at age for diff k, for each pop
# 2. use LSBlist to generate Jacobian matricies for simulations
# 3. get output from simulations ready for plotting
# 4. plot 1 - 4 panels, all pops together
# 5. plot 2 - 4 rows (k vals), columns are different peak spawning ages

```{r}
# --
# load functions, libraries
library(ggplot2)
library(gridExtra)
library(dplyr)
library(plyr)
library(ggrepel)
library(devtools)
library(broom)
source("C:/Users/provo/Documents/GitHub/popdy/cod_code/2_cod_functions.r") # load functions
source("C:/Users/provo/Documents/GitHub/popdy/cod_code/0_load_cod_data.r") # load cod data
source("C:/Users/provo/Documents/GitHub/popdy/cod_code/simulation_model_cod_v3.r") # load sim model
# read in eigentable - I'm using some information from the table
eigentable <- read.csv(file="C:/Users/provo/Documents/GitHub/popdy/cod_code/mikaelaLSB/eigentable.csv",header=T)
eigentable = as.data.frame(eigentable)

# ---
# prep output for new Leslie matricies
# Irish gives errors, remove from below
codNames <- c("Northsea","Coas","W_Baltic",
              "Faroe","NE_Arctic","Celtic",
              "Iceland","Kat","W_Scotland",
              "NGulf","GB","GM",
              "cod3NO","cod3M","cod2J3KL",
              "cod3Ps")
ks = c(0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9)
F.halfmax = 0 #for now, F is 0
LSBlist <- as.list(rep(NA,length(codNames)))
names(LSBlist) <- codNames
tknot =0

# ---
# this loop calculates LSB at age at different values of k

for (i in 1:length(LSBlist)) { # for each population
  # load parms for cod pop i
  source(file = paste('C:/Users/provo/Documents/GitHub/popdy/cod_pops/',names(LSBlist)[i], '.r', sep=''))
  # this should load parms: L_inf, K, TEMP, maxage, B0, B1
  
  # for each pop calc LSB at age for different k values
  lsb.at.k = matrix(0,nrow=maxage,ncol=length(ks)) #create empty matrix to store LSB at age at k
  for (u in 1:length(ks)) { # step through k values to create LSB at age
    lsb.at.k[,u] = calculate_LSB_at_age_by_F(data=datalist[[i]], maxage=maxage,
                                             L_inf=L_inf, K=K, TEMP=TEMP, F.halfmax=F.halfmax,
                                             B0=B0,B1=B1)
    lsb.at.k[,u] = lsb.at.k[,u]*ks[u] #multiply by k
  }
  LSBlist[[i]] = lsb.at.k #LSBlist should have length of codNames
}
rm(i,u)



# Goal: create list of arrays with Leslie matricies for each k-population
A3dlist = as.list(rep(NA,length(ks)))
for (i in 1:length(LSBlist)) {
  lsbatk = LSBlist[[i]]
  Alist = as.list(rep(NA,length(ks)))
  
  for (j in 1:length(ks)) { #this assumes I have LSB at the desired k values
    # start with empty matrix
    A = matrix(0,nrow=length(lsbatk[,1]),ncol=length(lsbatk[,1])) #create an empty 'leslie' matrix
    
    # create new Lelise matricies at each k
    A[1,]=lsbatk[,j]/sum(lsbatk[,j]) #fill in relative LSB at age across top row
    A[1,]=A[1,]*ks[j] #multiply relative fecundities by k
    
    for(a in 2:length(A[,1])-1){ #filling in survivals
      A[a+1,a]=1} #fill in survival of 1 on subdiagonal
    
    Alist[[j]] = A
    
  }
  A3d = array(unlist(Alist),dim=c(length(A[,1]),length(A[,1]),length(ks)))
  A3dlist[[i]] = A3d #store array 
}
rm(i,j)

# simulation using Leslie matricies at each k

timesteps = 1000
rm_first_timesteps = 100
alpha = 100
beta = 10000
initial_eggs = 1000
sig_r = 0.1
span.multiplier = 1 # what is this again?
```

The code below uses Jacobian matricies associated with each population and k slope to simulate timeseries of egg, 
recruits, and spawning adults. The matricies are stored in a list of 3D arrays.
```{r}
freq = seq(from=0.001111111, to=0.5, by=0.001111111) #all frequencies, for plotting later
output.norm.3dList = as.list(rep(NA,length(codNames)))
raw.3dList = as.list(rep(NA,length(codNames)))

# setting 'span' - a vector of odd integers to specify the smoothers
tmp <- ceiling(sqrt(length(1:(timesteps-rm_first_timesteps-1)))) #square root of timeseries length, rounded
if (tmp %% 2 == 0) {m <- tmp+1} else {m <- tmp} #make it odd, if the square root is even
m = m * span.multiplier
    
for (y in 1:length(A3dlist)) { #walk through each population
  A3d = A3dlist[[y]] #choose the 3d array of matricies, use these to simulate below
  
  outputL.raw = as.list(rep(NA,length(ks))) #create temp empty list for storing output at each k 
  outputL.norm = as.list(rep(NA,length(ks)))
                         
  for (k in 1:length(ks)) { #step through k values
    
    A = A3d[,,k] #choose the Leslie matrix for this k value
   
    output = sim_model(A=A, timesteps=timesteps, #simulate
                       alpha=alpha, beta=beta, 
                       sig_r=sig_r, initial_eggs=initial_eggs)  
    outputL.raw[[k]] = output #save sim output for each pop 
    
    # --- do the spectral analysis
    spec_all = matrix(NA,nrow=3,ncol=length(freq)) # matrix to store spec of egg, recruits, adults
    
    egg.sim = output[[2]][rm_first_timesteps:(timesteps-2)] #extract simulation of egg timeseries from model output
    spec_egg = spec.pgram(x=egg.sim, 
                          spans=c(m,m),plot = FALSE)
    spec_all[1,] <- spec_egg$spec #save spec output for plotting pops together, Helen Wearing says to multiply by 2
    
    rec.sim = output[[3]][rm_first_timesteps:(timesteps-2)] #extract sim of recruits
    spec_rec = spec.pgram(x=rec.sim, 
                          spans=c(m,m),plot = FALSE)
    spec_all[2,] = spec_rec$spec #save spec output for plotting pops together, Helen Wearing says to multiply by 2
    
    adu.sim = output[[4]][rm_first_timesteps:(timesteps-2)] #extract sim of spawning adults
    spec_adu = spec.pgram(x=adu.sim, 
                          spans=c(m,m),plot = FALSE)
    spec_all[3,] = spec_adu$spec #save spec output for plotting pops together, Helen Wearing says to multiply by 2
    
    outputL.norm[[k]] = spec_all
  }
  output.norm.3d = array(unlist(outputL.norm),dim=c(3,length(freq),length(ks)))
  output.norm.3dList[[y]] = output.norm.3d #store array 
  #output.raw.3d = array(unlist(outputL.raw), dim=c(3,timesteps,length(ks)))
  #output.raw.3dList[[y]] = output.raw.3d
}

# convert this into a dataframe
output.norm.3dList # for each pop, 3d object: 3 rows=eggs/rec/adu, col=spec output, stack=diff k 
```

The spec output from above is stored in a list of 3D arrays. The code below converts this list into a df.
```{r}
simpoplist = as.list(rep(NA,length(output.norm.3dList)))
names(simpoplist) <- codNames

for (e in 1:length(output.norm.3dList)) { #for each pop
  a = output.norm.3dList[[e]] #convert one pop array to df at a time
  
  simlist = as.list(rep(NA,length(ks))) #placeholder for storing spec output at k
  
  for (k in 1:length(ks)) { #for each k, convert sim output into df
  
    simout = as.data.frame(t(a[,,k])) #choose simout for each k
    names(simout) = c("eggs","recruits","adults")
    simout$freq = freq
    simlong = melt(simout,id.vars="freq")
    names(simlong) <- c("freq","out.type","value")
    simlong$kval <- rep(ks[k],length=length(freq))
    simlist[[k]] = simlong #store sim output df for each k
  } 
  df <- ldply(simlist, data.frame)
  df$codNames <- rep(codNames[e],length=length(df$freq))
  simpoplist[[e]] = df
}

spec.out.df <- ldply(simpoplist, data.frame)
table(spec.out.df$codNames)
```

Calculating the AUC.
```{r}
head(spec.out.df)
test <- spec.out.df[spec.out.df$out.type == "eggs" &
            spec.out.df$kval == "0.2" &
            spec.out.df$codNames == "Northsea",]
ggplot(test, aes(x=freq,y=value)) + geom_point()
sum(test$value * min(test$freq))

# Eggs
aucList <- as.list(NA,length=length(codNames))
for (i in 1:length(codNames)) {
  d <- spec.out.df[spec.out.df$codNames == codNames[i] & spec.out.df$out.type == "eggs",]
  auc <- rep(NA,length=length(ks))
  
  for (k in 1:length(ks)) {
    auc[k] <- sum( d[d$kval == ks[k],]$value * min(d$freq) )}
  
  aucdf <- data.frame(cbind(auc,ks,rep(codNames[i],length=length(ks))))
  names(aucdf) <- c("auc","kval","codNames")
  aucList[[i]] <- aucdf
}
auc.df.eggs <- ldply(aucList, data.frame)
rm(i,k,d,auc,aucdf,aucList)

# Recruits
aucList <- as.list(NA,length=length(codNames))
for (i in 1:length(codNames)) {
  d <- spec.out.df[spec.out.df$codNames == codNames[i] & spec.out.df$out.type == "recruits",]
  auc <- rep(NA,length=length(ks))
  
  for (k in 1:length(ks)) {
    auc[k] <- sum( d[d$kval == ks[k],]$value * min(d$freq) )}
  
  aucdf <- data.frame(cbind(auc,ks,rep(codNames[i],length=length(ks))))
  names(aucdf) <- c("auc","kval","codNames")
  aucList[[i]] <- aucdf
}
auc.df.recruits <- ldply(aucList, data.frame)
rm(i,k,d,auc,aucdf,aucList)

# Spawning adults
aucList <- as.list(NA,length=length(codNames))
for (i in 1:length(codNames)) {
  d <- spec.out.df[spec.out.df$codNames == codNames[i] & spec.out.df$out.type == "Nsize",]
  auc <- rep(NA,length=length(ks))
  
  for (k in 1:length(ks)) {
    auc[k] <- sum( d[d$kval == ks[k],]$value * min(d$freq) )}
  
  aucdf <- data.frame(cbind(auc,ks,rep(codNames[i],length=length(ks))))
  names(aucdf) <- c("auc","kval","codNames")
  aucList[[i]] <- aucdf
}
auc.df.adu <- ldply(aucList, data.frame)
rm(i,k,d,auc,aucdf,aucList)

```

```{r}

spec.out.df <- merge(spec.out.df,eigentable[c("codNames","temp")],all.x=TRUE)
spec.out.df <- merge(spec.out.df,eigentable[c("codNames","mode_age")],all.x=TRUE)
spec.out.df <- merge(spec.out.df,eigentable[c("codNames","sd_mode")],all.x=TRUE)
spec.out.df <- merge(spec.out.df,eigentable[c("codNames","max_ages")],all.x=TRUE)
spec.out.df <- merge(spec.out.df,eigentable[c("codNames","cvs_mode")],all.x=TRUE)


# Plot spectra
recruit.df <- spec.out.df[spec.out.df$out.type=="recruits" &
                            spec.out.df$kval %in% c("0.2","0.5","0.8"),]
recruit.df <- mutate(recruit.df, codNames = reorder(codNames, mode_age))
rp<- ggplot(recruit.df,
       aes(x=freq,y=value)) +
  geom_line() +
  facet_grid(kval~codNames) + ggtitle("recruits: increasing in peak spawning age")

# ---
egg.df <- spec.out.df[spec.out.df$out.type=="eggs" &
                            spec.out.df$kval %in% c("0.2","0.5","0.8"),]
egg.df <- mutate(egg.df, codNames = reorder(codNames, mode_age))
ep <- ggplot(egg.df,
       aes(x=freq,y=value)) +
  geom_line() +
  facet_grid(kval~codNames) + ggtitle("eggs: increasing in peak spawning age")
# ---
adu.df <- spec.out.df[spec.out.df$out.type=="adults" &
                            spec.out.df$kval %in% c("0.2","0.5","0.8"),]
ap <- ggplot(adu.df,
       aes(x=freq,y=value)) +
  geom_line() +
  facet_grid(kval~codNames) + ggtitle("spawning biomass")

#pdf(file='C:/Users/provo/Documents/GitHub/popdy/cod_figures/spectrum_plots_v1.pdf', width=20, height=7)
#rp
#ep
#dev.off()





```