---
title: "Jacobian and Leslie Simulations"
author: "Mikaela Provost"
date: "August 6, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# I. Load functions

## A) sim_model: simulates Leslie and Jacaobian matrices without density dependence

```{r}
sim_model <- function(A,timesteps,initial_eggs,maxage) {
  age_at_mat <- min(which(!A[1,] ==0)) # set inequality to min threshold for maturity, for cod it is always age 1
  ages = length(seq(1,maxage,by=1))
  N0 = c(initial_eggs, rep(0,ages-1)) #vector, length=num of ages, first age is initial eggs
  set.seed(2) #Set the seed so that every simulation uses same random sequence
  
  Nt = matrix(0,ages,timesteps) #empty matrix to store output: rows=ages, cols=time
  
  Nt[,1] = N0 #Put in initial values
  Nt[,2] = A %*% Nt[,1]  # multiply initial age vector with Jacobian to get 2nd age vector
  eggs = c(initial_eggs, rep(NA,timesteps-2)) #will save egg production here, this will be the input in BH
  eggs[2] = Nt[1,2]
  
  for(t in 1:(timesteps-2)){ #step through time
    #Save egg production, this is new number of age 1 individuals
    eggs[t+1] = Nt[1,t+1] 
    
    #perform population projection for one time step   
    Nt[,t+2] = A %*% Nt[,t+1] 
  }
  Nsize = colSums(Nt[age_at_mat:maxage,]) #sum rows that correspond to spawning adults
  eggs = eggs[1:(timesteps-1)] #egg production
  return(list(eggs=eggs, Nsize=Nsize))
}
```

## B) sim_model_dd: simulates Leslie and Jacaobian matrices with density dependence and noise

```{r}
sim_model_dd <- function(A,timesteps,alpha,beta,sig_r,initial_eggs) {
  age_at_mat <- min(which(!A[1,] ==0)) # set inequality to min threshold for maturity, for cod it is always age 1
  
  ages = length(seq(1,maxage,by=1))
  N0 = c(initial_eggs, rep(0,ages-1)) #vector, length=num of ages, first age is initial eggs
  set.seed(2) #Set the seed so that every simulation uses same random sequence
  
  Nt = matrix(0,ages,timesteps) #empty matrix to store output: rows=ages, cols=time
  
  Nt[,1] = N0 #Put in initial values
  Nt[,2] = A %*% Nt[,1]  # multiply initial age vector with Jacobian to get 2nd age vector
  eggs = c(initial_eggs, rep(NA,timesteps-2)) #will save egg production here, this will be the input in BH
  eggs[2] = Nt[1,2]
  recruits1 = eggs[1]/( (1/alpha) + (eggs[1]/beta) )
  recruits = c(0,rep(NA, timesteps-2)) #will save recruits here (output from BH)
  recruits[2] = recruits1 #we are ignoring recorinding recruits at time 0
  
  for(t in 1:(timesteps-2)){ #step through time
    #Save egg production, this is new number of age 1 individuals
    eggs[t+1] = Nt[1,t+1] 
    #save recruits, treat egg production as spawning biomass in BH
    recruits[t+1] = eggs[t]/( (1/alpha) + (eggs[t]/beta) ) #((alpha*eggs[t+1])/(1+beta*eggs[t+1])) 
    #replace age 1 in numbers-at-age matrix with recruits from BH, add noise    
    Nt[1,t+1] = (recruits[t+1])*exp(sig_r*rnorm(1,mean=0,sd=1)) 
    #perform population projection for one time step   
    Nt[,t+2] = A %*% Nt[,t+1] 
  }
  Nsize = colSums(Nt[age_at_mat:maxage,]) #sum rows that correspond to spawning adults
  eggs = eggs[1:(timesteps-1)] #egg production
  recruits = recruits[1:(timesteps-1)] #recruits produced via BH, before influence of environmental noise
  return(list(eggs=eggs, recruits=recruits, Nsize=Nsize))
}
```

## C) load the function: extract_first_eigen_value()

```{r warning=FALSE}
source("C:/Users/provo/Documents/GitHub/popdy/cod_code/2_cod_functions.r")

```

# II. Generate Leslie matrix

I use North Sea cod population as a test case. First I define all the North Sea parameters for assembling the Leslie matrix.

```{r}
#source(file = paste('C:/Users/provo/Documents/GitHub/popdy/cod_pops/','Northsea', '.r', sep=''))
# North Sea example
maxage = 10
K = 0.217
L_inf = 126
F.halfmax=0
B0 = -6.42
B1 = 1.72
tknot=0
MG = 0.427
```

I generate a Leslie matrix (A) and life table (NEAR) using the function assemble_Leslie().
```{r}
Leslieout = assemble_Leslie(maxage=maxage, K=K, L_inf=L_inf,F.halfmax=0, B0=B0, B1=B1, tknot=0)
NEAR = Leslieout$NEAR #Life table
LEP = sum(NEAR$egg_production) #sum egg production (not adjusted)
A = Leslieout$A #Leslie matrix 
A
```

# III. Simulate Leslie and Jacobian without density dependence

## A) Leslie simulation without density dependence

Before we simulate the Leslie matrix, we first have to standardize LEP so that LEP is constant among populations. Here I set LEP=1.1. To do this we adjust fecundities, fecundity-at-age should be: f-at-age/(LEP*0.9)

```{r}
conLEP = 1.1 #constant LEP
adjFec = round(1/conLEP,digits=1) #1/LEP is 1.1, which means alpha must be greater than this
A[1,] <- A[1,]/(LEP*adjFec) #adjusted Leslie matrix
```
Since the leading eigenvalue of the fecundity-adjusted-Leslie matrix is `r extract_first_eigen_value(A)`, we expect the population to exponentially grow because lambda1 > 1. The plot below shows population size increases exponentially.

```{r}
Leslie_output_withoutDD_lambda_small <- sim_model(A=A,timesteps=1000,initial_eggs=100,maxage=10)
plot(Leslie_output_withoutDD_lambda_small$Nsize,xlab="year",ylab="Pop size",main="Leslie simulation without\ndensity dependence (lambda1=1.02)")
```

## B) Jacobian simulation without density dependence (alpha=5.51)
Second, convert Leslie to Jacobian: multiply top row of Leslie by 1/(alpha x LEP^2). 
The slope at equilibrium is 1/(alpha x LEP^2) (see ch 4 in Loo's book or my appendix for derivation). 
Since we adjusted LEP to 1.1, we multiply to top row by: 1/(alpha*1.1^2) (we call this value 'k'). 
If alpha is 5.51 then k is `r round(1/(5.51*1.1^2),digits=2)`. 
```{r}
alpha = 5.51 #with this value of alpha, slope at equilibrium is 0.15
slope_at_equilibrium = 1/(alpha*conLEP^2) #we call the slope 'k'
A[1,] <- A[1,]*slope_at_equilibrium #convert to Jacobian
```

Lambda1 of the Jacobian is `r extract_first_eigen_value(A)`. Since lambda1 < 1 we expect a simulation of the Jacobian to converge to zero. The plot below shows the population is perturbed initially, but quickly goes back to zero. 
```{r}
Jacobian_output_withoutDD_lambda_small <- sim_model(A=A,timesteps=1000, initial_eggs=100,maxage=10)
plot(Jacobian_output_withoutDD_lambda_small$Nsize,xlab="year",ylab="Pop size",main="Jacobian simulation without\ndensity dependence (lambda1=0.76)")
```

## C) Jacobian simulation without density dependence (alpha=0.97)

If alpha is 0.97 then k is `r 1/(0.97*1.1^2)`. With this new Jacobian, the leading eigenvalue is `r extract_first_eigen_value(A)` which means we expect the population to approach zero but at a slower rate than before because lambda1 is closer to 1. 
```{r}
alpha0.97 = 0.97
A = Leslieout$A #reset Leslie matrix to original form without adjustments
A[1,] <- A[1,]/(LEP*adjFec) #adjust LEP in Leslie matrix to make LEP=1.1
slope_at_equilibrium = 1/(alpha0.97*conLEP^2) #we call the slope at equilibrium 'k'
A[1,] <- A[1,]*slope_at_equilibrium #convert Leslie to the Jacobian by multiplying top row 1/(alpha*LEP^2)
# Simulate Jacobian:
Jacobian_output_withoutDD_lambda_big <- sim_model(A=A,timesteps=1000, initial_eggs=100,maxage=10)
plot(Jacobian_output_withoutDD_lambda_big$Nsize,xlab="year",ylab="Pop size",main="Jacobian simulation without\ndensity dependence (lambda1=0.99)")
```

# IV. Simulate Leslie and Jacobian with density dependence and noise

## A) Simulate Leslie with density dependence (alpha = 5.51):
```{r}
# First, set up the fecundity-adjusted Leslie matrix:
A = Leslieout$A #reset Leslie matrix to original form without adjustments
A[1,] <- A[1,]/(LEP*adjFec) #adjust LEP in Leslie matrix to make LEP=1.1
```

Lambda1 of the Leslie matrix is `r extract_first_eigen_value(A)`. A plot of annual recruitment before noise is below.

```{r}
alpha5.51 <- 5.51 # define alpha for BH in DD
Leslie_output_withDD_lambda_small <- sim_model_dd(A=A,timesteps=1000,alpha=5.51,beta=1000,sig_r=0.3,initial_eggs=100)
plot(Leslie_output_withDD_lambda_small$recruits,xlab="year",ylab="recruits (before noise)",main="Leslie simulation with\ndensity dependence (lambda1=1.05)",type="l")
```

## B) Simulate Jacobian with density dependence (alpha = 5.51):
```{r}
# Second, set up the Jacobian:
slope_at_equilibrium = 1/(alpha5.51*conLEP^2) #we call the slope at equilibrium 'k'
A[1,] <- A[1,]*slope_at_equilibrium #convert Leslie to the Jacobian by multiplying top row 1/(alpha*LEP^2)
# Let's check the leading eigenvalue of this Jacobian:
```

Lambda1 of the Jacobian is `r extract_first_eigen_value(A)`. Below is a plot showing a simulation of annual recruits (before noise) using the Jacobian maxtric.

```{r}
Jacobian_output_withDD_lambda1_small <- sim_model_dd(A=A,timesteps=1000,alpha=5.51,beta=1000,sig_r=0.3,initial_eggs=100)
plot(Jacobian_output_withDD_lambda1_small$recruits,xlab="year",ylab="recruits (before noise)",main="Jacobian simulation with\ndensity dependence (lambda1=0.77)",type="l")
```

## C) Simulate Leslie with density dependence (alpha = 0.97)

Set up the Leslie matrix:
```{r}
alpha0.97 <- 0.97
A = Leslieout$A #reset Leslie matrix to original form without adjustments
A[1,] <- A[1,]/(LEP*adjFec) #adjust LEP in Leslie matrix to make LEP=1.1
```

Lambda1 of the Leslie is `r extract_first_eigen_value(A)`. Below is a plot showing a simulation of annual recruits (before noise) using the Leslie matrix. 

```{r}
# Simulate adjusted Leslie with density dependence:
Leslie_output_withDD_lambda_big <- sim_model_dd(A=A,timesteps=1000,alpha=0.97,beta=1000,sig_r=0.3,initial_eggs=100)
plot(Leslie_output_withDD_lambda_big$recruits,xlab="year",ylab="recruits (before noise)",main="Leslie simulation with\ndensity dependence (lambda1=1.05)",type="l")
```

## D) Simulate Jacobian with density dependence (alpha = 0.97)
```{r}
# Next, set up the Jacobian:
slope_at_equilibrium = 1/(alpha0.97*conLEP^2) #we call the slope at equilibrium 'k'
A[1,] <- A[1,]*slope_at_equilibrium #convert Leslie to the Jacobian 
```

Lambda1 of the Jacobian is `r extract_first_eigen_value(A)`. Below is a plot showing a simulation of annual recruits (before noise) using the Jacobian matrix. 

```{r}
# Simulate model w/dd:
Jacobian_output_withDD_lambda1_big <- sim_model_dd(A=A,timesteps=1000,alpha=0.97,beta=1000,sig_r=0.3,initial_eggs=100)
plot(Jacobian_output_withDD_lambda1_big$recruits,xlab="year",ylab="recruits (before noise)",main="Jacobian simulation with\ndensity dependence (lambda1=0.99)",type="l")

```