---
title: "parms"
author: "Mikaela Provost"
date: "March 26, 2019"
output: html_document
---

```{r setup, include=FALSE}
# load packages
```

## Gathering parameters for PMFC species

I'm gathering parameters for XX species which are managed by the Pacific Fisheries Management Council (see <https://www.pcouncil.org/>). A lot of the life history information I need is stored on Fishbase. Instead of clicking on each species webpage and recording life history values I'm going to webscrape Fishbase website to record this information. 

```{r echo=FALSE, results='asis'}
Common name <- c("Pacific Sardine","Pacific Mackerel")
library(knitr)
kable(inserttablehere, caption = "table caption here.")
```

see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
